MRP Case Study

### 2020-03-20 Code/mrp/Case Study .Rmd
This is a working hierarchal multinomial logit model. I am working on making it more interpretable. This comes from the ART forum 2017.

```{r message = FALSE}
library(tidyverse) #Load packages
library(rstan)
library(rstanarm)
library(ggplot2)
library(bayesplot)
theme_set(bayesplot::theme_default())
library(dplyr)
library(tidyr)
library(tidybayes)

options(mc.cores = parallel::detectCores()) # Set Stan to use all availible cores
rstan_options(auto_write = TRUE) # Don't recompile Stan code that hasn't changed
```

@@

```{r}
simulate_mrp_data <- function(n) {
  J <- c(2, 3, 7, 3, 50) # male or not, eth, age, income level, state
  poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) # Columns of post-strat matrix, plus one for size
  colnames(poststrat) <- c("male", "eth", "age","income", "state",'N')
  count <- 0
  for (i1 in 1:J[1]){
    for (i2 in 1:J[2]){
      for (i3 in 1:J[3]){
        for (i4 in 1:J[4]){
          for (i5 in 1:J[5]){
              count <- count + 1
              # Fill them in so we know what category we are referring to
              poststrat[count, 1:5] <- c(i1-1, i2, i3, i4, i5)
          }
        }
      }
    }
  }
  # Proportion in each sample in the population
  p_male <- c(0.52, 0.48)
  p_eth <- c(0.5, 0.2, 0.3)
  p_age <- c(0.2,.1,0.2,0.2, 0.10, 0.1, 0.1)
  p_income<-c(.50,.35,.15)
  p_state_tmp<-runif(50,10,20)
  p_state<-p_state_tmp/sum(p_state_tmp)
  poststrat$N<-0
  for (j in 1:prod(J)){
    poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
      p_age[poststrat[j,3]]*p_income[poststrat[j,4]]*p_state[poststrat[j,5]]) #Adjust the N to be the number observed in each category in each group
  }

  # Now let's adjust for the probability of response
  p_response_baseline <- 0.01
  p_response_male <- c(2, 0.8) / 2.8
  p_response_eth <- c(1, 1.2, 2.5) / 4.7
  p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
  p_response_inc <- c(1, 0.9, 0.8) / 2.7
  p_response_state <- rbeta(50, 1, 1)
  p_response_state <- p_response_state / sum(p_response_state)
  p_response <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    p_response[j] <-
      p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
      p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]] *
      p_response_inc[poststrat[j, 4]] * p_response_state[poststrat[j, 5]]
  }
  people <- sample(prod(J), n, replace = TRUE, prob = poststrat$N * p_response)

  ## For respondent i, people[i] is that person's poststrat cell,
  ## some number between 1 and 32
  n_cell <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    n_cell[j] <- sum(people == j)
  }

  coef_male <- c(0,-0.3)
  coef_eth <- c(0, 0.6, 0.9)
  coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
  coef_income <- c(0,-0.2, 0.6)
  coef_state <- c(0, round(rnorm(49, 0, 1), 1))
  coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
                           c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
  true_popn <- data.frame(poststrat[, 1:5], service_failure = rep(NA, prod(J)))
  for (j in 1:prod(J)) {
    true_popn$satisfaction[j] <- plogis(
      coef_male[poststrat[j, 1] + 1] +
        coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
        coef_income[poststrat[j, 4]] + coef_state[poststrat[j, 5]] +
        coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
      )
  }

  #male or not, eth, age, income level, state, city
  y <- rbinom(n, 1, true_popn$satisfaction[people])
  male <- poststrat[people, 1]
  eth <- poststrat[people, 2]
  age <- poststrat[people, 3]
  income <- poststrat[people, 4]
  state <- poststrat[people, 5]

  sample <- data.frame(service_failure = y,
                       male, age, eth, income, state,
                       id = 1:length(people))

  #Make all numeric:
  for (i in 1:ncol(poststrat)) {
    poststrat[, i] <- as.numeric(poststrat[, i])
  }
  for (i in 1:ncol(true_popn)) {
    true_popn[, i] <- as.numeric(true_popn[, i])
  }
  for (i in 1:ncol(sample)) {
    sample[, i] <- as.numeric(sample[, i])
  }
  list(
    sample = sample,
    poststrat = poststrat,
    true_popn = true_popn
  )
}
```

```{r}
mrp_sim <- simulate_mrp_data(n=1200)
sample <- mrp_sim[["sample"]]

print(sample)
```

```{r echo=FALSE}

J <- c(2, 3, 7, 3, 50) # male or not, eth, age, income level, state
  poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) # Columns of post-strat matrix, plus one for size
  colnames(poststrat) <- c("male", "eth", "age","income", "state",'N')
  count <- 0
  for (i1 in 1:J[1]){
    for (i2 in 1:J[2]){
      for (i3 in 1:J[3]){
        for (i4 in 1:J[4]){
          for (i5 in 1:J[5]){
              count <- count + 1
              # Fill them in so we know what category we are referring to
              poststrat[count, 1:5] <- c(i1-1, i2, i3, i4, i5)
          }
        }
      }
    }
  }
  # Proportion in each sample in the population
  p_male <- c(0.52, 0.48)
  p_eth <- c(0.5, 0.2, 0.3)
  p_age <- c(0.2,.1,0.2,0.2, 0.10, 0.1, 0.1)
  p_income<-c(.50,.35,.15)
  p_state_tmp<-runif(50,10,20)
  p_state<-p_state_tmp/sum(p_state_tmp)
  poststrat$N<-0
  for (j in 1:prod(J)){
    poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
      p_age[poststrat[j,3]]*p_income[poststrat[j,4]]*p_state[poststrat[j,5]]) #Adjust the N to be the number observed in each category in each group
  }

  # Now let's adjust for the probability of response
  p_response_baseline <- 0.01
  p_response_male <- c(2, 0.8) / 2.8
  p_response_eth <- c(1, 1.2, 2.5) / 4.7
  p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
  p_response_inc <- c(1, 0.9, 0.8) / 2.7
  p_response_state <- rbeta(50, 1, 1)
  p_response_state <- p_response_state / sum(p_response_state)
  p_response <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    p_response[j] <-
      p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
      p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]] *
      p_response_inc[poststrat[j, 4]] * p_response_state[poststrat[j, 5]]
  }

  coef_male <- c(0,-0.3)
  coef_eth <- c(0, 0.6, 0.9)
  coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
  coef_income <- c(0,-0.2, 0.6)
  coef_state <- c(0, round(rnorm(49, 0, 1), 1))
  coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
                           c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
  true_popn <- data.frame(poststrat[, 1:5], service_failure = rep(NA, prod(J)))
  for (j in 1:prod(J)) {
    true_popn$satisfaction[j] <- plogis(
      coef_male[poststrat[j, 1] + 1] +
        coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
        coef_income[poststrat[j, 4]] + coef_state[poststrat[j, 5]] +
        coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
      )
  }
```

Now that we have our simulated data, we now need our hierarchal stan model.

@@

// Index values, observations, and covariates.
data {
  int<lower = 1> N;                      // Number of observations.
  int<lower = 1> K;                      // Number of groups.
  int<lower = 1> I;                      // Number of observation-level covariates.

  real satisfaction[N];                        // Vector of observations.
  int<lower = 1, upper = K> male[N];        // Vector of group assignments.
  int eth[N];                               // Vector of ethnicity covariates.
  int age[N];                               // Vector of age covariates.
  int income[N];                          // Vector of income covariates.
  int state[N];                          // Vector of state covariates.

  real gamma_mean;                       // Mean for the hyperprior on gamma.
  real<lower = 0> gamma_var;             // Variance for the hyperprior on gamma.
  real<lower = 0> tau_min;               // Minimum for the hyperprior on tau.
  real<lower = 0> tau_max;               // Maximum for the hyperprior on tau.
  real<lower = 0> sigma_min;             // Minimum for the hyperprior on tau.
  real<lower = 0> sigma_max;             // Maximum for the hyperprior on tau.
}

// Parameters and hyperparameters.
parameters {
  matrix[K, (I - 1)] alpha;              // Matrix of observation-level brand coefficients.
  vector[K] beta;                        // Vector of observation-level price coefficients.
  real gamma;                            // Mean of the population model.
  real<lower=0> tau;                     // Variance of the population model.
  real<lower=0> sigma;                   // Variance of the observation model.
}

// Hierarchical regression.
model {
  // Declare mu for use in the linear model.
  vector[N] mu;

  // Hyperpriors and prior.
  gamma ~ normal(gamma_mean, gamma_var);
  tau ~ uniform(tau_min, tau_max);
  sigma ~ uniform(sigma_min, sigma_max);

  // Population model and likelihood.
  for (k in 1:K) {
    alpha[k,] ~ normal(gamma, tau);
    beta[k] ~ normal(gamma, tau);
  }
  for (n in 1:N) {
    mu[n] = alpha[male[n], eth[n]] + beta[male[n]] * age[n] + beta[male[n]] * 
    income[n] + beta[male[n]] * state[n];
  }
  cat_pref ~ normal(mu, sigma);
}

// Generate predictions using the posterior.
generated quantities {
  vector[N] mu_pc;                       // Declare mu for predicted linear model.
  real cat_pref_pc[N];                     // Vector of predicted observations.

  // Generate posterior prediction distribution.
  for (n in 1:N) {
    mu_pc[n] = alpha[male[n], eth[n]] + beta[male[n]] * age[n] + beta[male[n]] * 
    income[n] + beta[male[n]] * state[n];
    cat_pref_pc[n] = normal_rng(mu_pc[n], sigma);
  }
}

@@

As we will also need to do some prior and posterior checks, and for that we will need the generative model.

// Index values, observations, covariates, and hyperior values.
data {
  int<lower = 1> N;                      // Number of observations.
  int<lower = 1> K;                      // Number of groups.
  int<lower = 1> I;                      // Number of observation-level covariates.

  int<lower = 1, upper = K> male[N];        // Vector of group assignments.
  int eth[N];                               // Vector of ethnicity covariates.
  int age[N];                               // Vector of age covariates.
  int income[N];                          // Vector of income covariates.
  int state[N];                          // Vector of state covariates.

  real gamma_mean;                       // Mean for the hyperprior on gamma.
  real<lower = 0> gamma_var;             // Variance for the hyperprior on gamma.
  real<lower = 0> tau_min;               // Minimum for the hyperprior on tau.
  real<lower = 0> tau_max;               // Maximum for the hyperprior on tau.
  real<lower = 0> sigma_min;             // Minimum for the hyperprior on tau.
  real<lower = 0> sigma_max;             // Maximum for the hyperprior on tau.
}

// Generate data according to the hierarchical regression.
generated quantities {
  matrix[K, (I - 1)] alpha;              // Matrix of observation-level brand coefficients.
  vector[K] beta;                        // Vector of observation-level price coefficients.
  real gamma;                            // Mean of the population model.
  real<lower=0> tau;                     // Variance of the population model.
  real<lower=0> sigma;                   // Variance of the observation model.

  vector[N] mu;                          // Declare mu for linear model.
  real cat_pref[N];                        // Vector of observations.

  gamma = normal_rng(gamma_mean, gamma_var);
  tau = uniform_rng(tau_min, tau_max);
  sigma = uniform_rng(sigma_min, sigma_max);

  // Draw parameter values and generate data.
  for (k in 1:K) {
    for (i in 1:(I - 1)) {
      alpha[k, i] = normal_rng(gamma, tau);
    }
    beta[k] = normal_rng(gamma, tau);
  }
  for (n in 1:N) {
    mu[n] = alpha[male[n], eth[n]] + beta[male[n]] * age[n] + beta[male[n]] * 
    income[n] + beta[male[n]] * state[n];
    cat_pref[n] = normal_rng(mu[n], sigma);
  }
}

@@

```{r}
sim_values <- list(
  N = 6300,                                       # Number of observations.
  K = 2,                                         # Number of groups.
  I = 65,                                         # Number of observation-level covariates.

  male = sample(2, 6300, replace = TRUE),            # Vector of group assignments.
  eth = sample(3, 6300, replace = TRUE),       # Vector of brands covariates.
  age = sample(7, 6300, replace = TRUE),        # Vector of price covariates.
  income = sample(3, 6300, replace = TRUE),
  state = sample(50, 6300, replace = TRUE),

  gamma_mean = 0.5,                              # Mean for the hyperprior on gamma.
  gamma_var = 0.2,                               # Variance for the hyperprior on gamma.
  tau_min = 0,                                   # Minimum for the hyperprior on tau.
  tau_max = .1,                                   # Maximum for the hyperprior on tau.
  sigma_min = 0,                                 # Minimum for the hyperprior on tau.
  sigma_max = .1                                  # Maximum for the hyperprior on tau.
)
# Generate data.
sim_data <- stan(
  file = here::here("Code/mrp/generate_data.stan"),
  data = sim_values,
  iter = 10,
  chains = 1,
  seed = 42,
  algorithm = "Fixed_param"
)
```

```{r}
# Extract the simulated data.
prior_pc <- tibble(
  cat_pref = as.vector(extract(sim_data)$cat_pref)
)
# Plot the prior predictive distribution.
prior_pc %>% 
  ggplot(aes(x = cat_pref)) +
  geom_density()
```

```{r}
# # Specify data.
# data <- list(
#   N = nrow(true_popn),                     # Number of observations.
#   K = 2,                    # Number of groups.
#   I = 65,                                # Number of observation-level covariates.
#   
#   cat_pref = true_popn$cat_pref,               # Vector of observations.
#   male = as.numeric(as.factor(true_popn$male)),            # Vector of group assignments.
#   eth = true_popn$eth,       # Vector of brands covariates.
#   age = true_popn$age,        # Vector of price covariates.
#   income = true_popn$income,
#   state = true_popn$state,
#   
#   gamma_mean = 0.5,                                # Mean for the hyperprior on gamma.
#   gamma_var = .2,                                 # Variance for the hyperprior on gamma.
#   tau_min = 0,                                   # Minimum for the hyperprior on tau.
#   tau_max = 1,                                   # Maximum for the hyperprior on tau.
#   sigma_min = 0,                                 # Minimum for the hyperprior on tau.
#   sigma_max = 1                                 # Maximum for the hyperprior on tau.
# )
# # Calibrate the model.
# model03 <- stan(
#   file = here::here("Code/mrp/hierarchical_model.stan"),
#   data = data,
#   control = list(adapt_delta = 0.99),
#   seed = 42
# )
```

```{r}
# Posterior predictive check.
# post_pc03 <- tibble(
#   # Extract the posterior predicted values.
#   intent = as.vector(extract(model03)$cat_pref_pc)
# )
# # Plot the posterior predictive distribution.
# ggplot(post_pc03, aes(x = intent)) +
#   geom_histogram() +
#   xlim(1, 10) 
# ggplot(purchase_intent, aes(x = intent)) +
#   geom_histogram() +
#   xlim(1, 10) 
```

```{r}
# model03 %>%
#   gather_draws(alpha[n, i]) %>%
#   unite(.variable, .variable, n, i) %>%
#   ggplot(aes(x = .value, y = .variable)) +
#   geom_halfeyeh(.width = .95) +
#   facet_wrap(
#     ~ .variable,
#     nrow = data$K,
#     ncol = (data$I - 1),
#     scales = "free"
#   )
```

```{r}
# Plot the betas.
# model03 %>%
#   gather_draws(beta[i]) %>%
#   unite(.variable, .variable, i) %>%
#   ggplot(aes(x = .value, y = .variable)) +
#   geom_halfeyeh(.width = .95) +
#   facet_wrap(
#     ~ .variable,
#     nrow = data$K,
#     ncol = 1,
#     scales = "free"
#   )
```

### Postratification 
Note: for now I have set these chuncks to eval = false because we don
t have a poststrat_prob yet so it freaks out

For the postratification portion we take the estimate that we get from the model times poststrat$/N / sum(postsrat$N)

```{r eval = FALSE}
poststrat_prob <- posterior_prob %*% poststrat$N / sum(poststrat$N)
model_popn_pref <- c(mean = mean(poststrat_prob), sd = sd(poststrat_prob))
round(model_popn_pref, 3)
```

Because we are using simulated data we can see how our poststratified predictions compare to the true population

```{r eval = FALSE}
sample_popn_pref <- mean(sample$satisfaction)
round(sample_popn_pref, 3)

compare2 <- compare2 +
  geom_hline(yintercept = model_popn_pref[1], colour = '#2ca25f', size = 1) +
  geom_text(aes(x = 5.2, y = model_popn_pref[1] + .025), label = "MRP", colour = '#2ca25f')
bayesplot_grid(compare, compare2,
               grid_args = list(nrow = 1, widths = c(8, 1)))
```

>>>>

```{r message = FALSE}

library(tidyverse) #Load packages
library(rstan)
library(rstanarm)
library(ggplot2)
library(bayesplot)
theme_set(bayesplot::theme_default())
library(dplyr)
library(tidyr)
library(tidybayes)
options(mc.cores = parallel::detectCores()) # Set Stan to use all availible cores
rstan_options(auto_write = TRUE) # Don't recompile Stan code that hasn't changed
```

@@

```{r}
simulate_mrp_data <- function(n) {
  J <- c(2, 3, 7, 3, 50) # male or not, eth, age, income level, state
  poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) # Columns of post-strat matrix, plus one for size
  colnames(poststrat) <- c("male", "eth", "age","income", "state",'N')
  count <- 0
  for (i1 in 1:J[1]){
    for (i2 in 1:J[2]){
      for (i3 in 1:J[3]){
        for (i4 in 1:J[4]){
          for (i5 in 1:J[5]){
              count <- count + 1
              # Fill them in so we know what category we are referring to
              poststrat[count, 1:5] <- c(i1-1, i2, i3, i4, i5)
          }
        }
      }
    }
  }
  # Proportion in each sample in the population
  p_male <- c(0.52, 0.48)
  p_eth <- c(0.5, 0.2, 0.3)
  p_age <- c(0.2,.1,0.2,0.2, 0.10, 0.1, 0.1)
  p_income<-c(.50,.35,.15)
  p_state_tmp<-runif(50,10,20)
  p_state<-p_state_tmp/sum(p_state_tmp)
  poststrat$N<-0
  for (j in 1:prod(J)){
    poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
      p_age[poststrat[j,3]]*p_income[poststrat[j,4]]*p_state[poststrat[j,5]]) #Adjust the N to be the number observed in each category in each group
  }
  # Now let's adjust for the probability of response
  p_response_baseline <- 0.01
  p_response_male <- c(2, 0.8) / 2.8
  p_response_eth <- c(1, 1.2, 2.5) / 4.7
  p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
  p_response_inc <- c(1, 0.9, 0.8) / 2.7
  p_response_state <- rbeta(50, 1, 1)
  p_response_state <- p_response_state / sum(p_response_state)
  p_response <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    p_response[j] <-
      p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
      p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]] *
      p_response_inc[poststrat[j, 4]] * p_response_state[poststrat[j, 5]]
  }
  people <- sample(prod(J), n, replace = TRUE, prob = poststrat$N * p_response)
  ## For respondent i, people[i] is that person's poststrat cell,
  ## some number between 1 and 32
  n_cell <- rep(NA, prod(J))
  for (j in 1:prod(J)) {
    n_cell[j] <- sum(people == j)
  }
  coef_male <- c(0,-0.3)
  coef_eth <- c(0, 0.6, 0.9)
  coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
  coef_income <- c(0,-0.2, 0.6)
  coef_state <- c(0, round(rnorm(49, 0, 1), 1))
  coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
                           c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
  true_popn <- data.frame(poststrat[, 1:5], service_failure = rep(NA, prod(J)))
  for (j in 1:prod(J)) {
    true_popn$satisfaction[j] <- plogis(
      coef_male[poststrat[j, 1] + 1] +
        coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
        coef_income[poststrat[j, 4]] + coef_state[poststrat[j, 5]] +
        coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
      )
  }
  #male or not, eth, age, income level, state, city
  y <- rbinom(n, 1, true_popn$satisfaction[people])
  male <- poststrat[people, 1]
  eth <- poststrat[people, 2]
  age <- poststrat[people, 3]
  income <- poststrat[people, 4]
  state <- poststrat[people, 5]
  sample <- data.frame(service_failure = y,
                       male, age, eth, income, state,
                       id = 1:length(people))
  #Make all numeric:
  for (i in 1:ncol(poststrat)) {
    poststrat[, i] <- as.numeric(poststrat[, i])
  }
  for (i in 1:ncol(true_popn)) {
    true_popn[, i] <- as.numeric(true_popn[, i])
  }
  for (i in 1:ncol(sample)) {
    sample[, i] <- as.numeric(sample[, i])
  }
  list(
    sample = sample,
    poststrat = poststrat,
    true_popn = true_popn
  )
}
```

```{r}
mrp_sim <- simulate_mrp_data(n=1200)
sample <- mrp_sim[["sample"]]
poststrat <- mrp_sim[["poststrat"]]
true_popn <- mrp_sim[["true_popn"]]
head(sample)
```

```{r}
stan_data <- list( 
  N = nrow(sample), 
  service_failure = sample$service_failure,
  male = sample$male, 
  age = sample$age,
  eth = sample$eth, 
  income = sample$income, 
  state = sample$state)
```

We now have need for a logistic regresson model in order to predict a binary outcome for 'service_failure' so we will make a hierarchal logit model

# Hierarchal Logit Model

@@

data {
  int<lower=2> C; // # of alternatives (choices) in each scenario
  int<lower=1> K; // # of covariates of alternatives
  int<lower=1> R; // # of respondents
  int<lower=1> S; // # of scenarios per respondent
  int<lower=0> G; // # of respondent covariates 
  int<lower=1,upper=C> Y[R, S]; // observed choices
  matrix[C, K] X[R, S]; // matrix of attributes for each obs
  matrix[G, R] Z; // vector of covariates for each respondent
}

parameters {
  matrix[K, R] Beta;
  matrix[K, G] Theta;
  corr_matrix[K] Omega;
  vector<lower=0>[K] tau;
}

transformed parameters {
  cov_matrix[K] Sigma = quad_form_diag(Omega, tau);
}

model {
  //priors
  to_vector(Theta) ~ normal(0, 10);
  tau ~ cauchy(0, 2.5); 
  Omega ~ lkj_corr(2);
  //likelihood
  for (r in 1:R) {
    Beta[,r] ~ multi_normal(Theta*Z[,r], Sigma);	
    for (s in 1:S)
      Y[r,s] ~ categorical_logit(X[r,s]*Beta[,r]);
  }
}

@@

```{r}
# ========== Load libraries and set options ==========
library(rstan)
library(MASS)
library(shinystan)

# writes a compiled Stan program to the disk to avoid recompiling
rstan_options(auto_write=TRUE) 
# allows Stan chains to run in parallel on multiprocessor machines
options(mc.cores = parallel::detectCores())

# ========== Test Stan with synthetic data ============

# function to generate mnl data
generate_hmnl_data <- function(R=100, S=30, C=3, 
                               Theta=matrix(rep(1, 8), nrow=2), 
                               Sigma=diag(0.1, 4)){
  K <- ncol(Theta)
  G <- nrow(Theta)
  Y <- array(dim=c(R, S))
  X <- array(rnorm(R*S*C*K), dim=c(R, S, C, K)) # normal covariates
  Z <- array(dim=c(G, R))
  Z[1,] <- 1  # intercept
  if (G > 1) {
    Z[2:G,] <- rnorm(R*(G-1)) # normal covariates
  }
  Beta <- array(dim=c(K, R))
  for (r in 1:R) {
    Beta[,r] <- mvrnorm(n=1, mu=Z[,r]%*%Theta, Sigma=Sigma)
    for (s in 1:S)
      Y[r,s] <- sample(x=C, size=1, prob=exp(X[r,s,,]%*%Beta[,r])) # logit formula
   }
  list(R=R, S=S, C=C, K=K, G=G, Y=Y, X=X, Z=Z, 
       beta.true=beta, Theta.true=Theta, Sigma.true=Sigma)
}

d1 <- generate_hmnl_data()
str(d1)

test.stan <- stan(file="hmnl.stan", data=d1, iter=1000, chains=4)

plot(test.stan, plotfun="trace", pars=("Theta"))

plot(test.stan, plotfun="trace", pars=c("tau"))

plot(test.stan, plotfun="trace", pars=("Omega"))

plot(test.stan, pars=c("Theta", "tau", "Omega"))

plot(test.stan, pars=c("Theta", "Sigma"))

summary(test.stan, pars=c("Theta"))$summary

summary(test.stan, pars=c("tau"))$summary

summary(test.stan, pars=c("Omega"))$summary
```

```{r}
# ========= Read in Chocolate Data and Prep for Stan ==========
rm(list=ls()) # tidy up
choc.df <- read.csv("cbc_chocolate.csv")

# Coding the chocolate data (this ought to be a function)
choc.contrasts <- list(Brand = "contr.sum", Type = "contr.sum")
choc.coded <- model.matrix(~ Brand + Type, data = choc.df, contrasts = choc.contrasts)
choc.coded <- choc.coded[,2:ncol(choc.coded)] # remove intercept
# Fix the bad labels from contr.sum
choc.names <- c("BrandDove", "BrandGhirardelli", "BrandGodiva", "BrandHersheys", 
                "TypeDark", "TypeDarkNuts", "TypeMilk", "TypeMilkNuts")
colnames(choc.coded) <- choc.names
choc.df <- cbind(choc.df, choc.coded)

head(choc.df)

# Munge into Stan list
R <- length(unique(choc.df$Ind))
S <- length(unique(choc.df$Trial))
C <- max(choc.df$Alt)
K <- 9
Y <- array(dim=c(R, S))
X <- array(rnorm(R*S*C*K), dim=c(R, S, C, K)) 
Z <- array(1, dim=c(1, R)) # intercept only
for (r in 1:R) { # respondents
  for (s in 1:S){ # choice scenarios
    scenario <- choc.df[choc.df$Ind==unique(choc.df$Ind)[r] & 
                        choc.df$Trial==unique(choc.df$Trial)[s], ]
    X[r,s,,] <- data.matrix(scenario[,c(7, 9:16)]) # price and coded brand and type
    Y[r,s] <- scenario$Alt[as.logical(scenario$Chosen)]
  }
}

choc.standata <- list(C=C, K=K, R=R, S=S, G=1, Y=Y, X=X, Z=Z)
str(choc.standata)

rm(Y, X, Z, R, S, r, s, C, K, choc.contrasts, scenario, choc.coded)

@@

```{r}

# ========== Run Stan and Check Convergence =========
choc.stan <- stan(file="hmnl.stan", data=choc.standata)

plot(choc.stan, plotfun="trace", pars=("Theta"))

plot(choc.stan, plotfun="trace", pars=c("tau"))

plot(choc.stan, plotfun="trace", pars=c("Omega[1,2]"))

plot(choc.stan, plotfun="trace", pars=paste("Beta[", 1:9, ",1]", sep="")) # resp 1

summary(choc.stan)$summary[,c("Rhat", "n_eff")]

# Convergence checking gets tedious with a lot of parameters, so automate
check_fit <- function(fit) {
  summ <- summary(fit)$summary
  range_rhat <- range(summ[ , 'Rhat'])
  rhat_ok <- 0.99 <= range_rhat[1] && range_rhat[2] <= 1.1
  range_neff <- range(summ[ , 'n_eff'])
  neff_ok <- range_neff[1] >= 400
  sp <- rstan::get_sampler_params(fit, inc_warmup=FALSE)
  max_divergent <- max(sapply(sp, function(p){ sum(p[ , 'divergent__']) }))
  no_divergent <- max_divergent == 0

  list(ok = rhat_ok && neff_ok && no_divergent,
       range_rhat = range_rhat,
       range_neff = range_neff,
       max_divergent = max_divergent)
}

check_fit(choc.stan)
```

```{r}
# ========== Summarize Posterior ==========
choc.names
plot(choc.stan, pars=c("Theta", "tau"))

plot(choc.stan, pars=c("Omega"))

plot(choc.stan, pars=paste("Beta[", 1:9, ",1]", sep="")) + ggtitle("Respondent 1: Likes Milk Chocolate")

plot(choc.stan, pars=paste("Beta[", 1:9, ",2]", sep="")) + ggtitle("Respondent 2: Likes Dark Chocolate")

launch_shinystan(choc.stan)

# ========== Simulate Shares from Model =========
# function for computing shares from beta.draws for an hmnl model
shares.hmnl.post <- function(beta.draws, X) # X is attribute matrix for scenario
{
  R <- dim(beta.draws)[3]  # respondents
  D <- dim(beta.draws)[1]  # draws
  shares <- array(NA, dim=c(nrow(X), R, D))
  for (d in 1:D) {
    for (r in 1:R) {
      beta <- beta.draws[d,,r] 
      V <- exp(X %*% beta)  
      shares[,r,d] <- V/sum(V)
    }
  }
  shares
}

choc.standata$X[1,1,,] # note option 2 is dark
shares <- shares.hmnl.post(extract(choc.stan, pars=c("Beta"))$Beta, 
                          choc.standata$X[1,1,,])
str(shares)

apply(shares, 1, quantile, probs=c(0.5, 0.025, 0.975))
apply(shares, 1:2, quantile, probs=c(0.5, 0.025, 0.975))
